### 2019-03-01 langs-figure.tex

\documentclass[8pt,letterpaper]{article}

\usepackage{pslatex}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{amssymb}

\usepackage{tikz-dependency}
\usepackage{longtable}

\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\E}[0]{\mathbb{E}}
\newcommand{\Ff}[0]{\mathcal{F}}

\usepackage{multirow}

\newcommand{\soft}[1]{}
\newcommand{\nopreview}[1]{}
\newcommand\comment[1]{{\color{red}#1}}
\newcommand\mhahn[1]{{\color{red}(#1)}}

\usepackage{amsthm}

\newcommand{\thetad}[0]{{\theta_d}}
\newcommand{\thetal}[0]{{\theta_{LM}}}

\newcounter{theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{question}[theorem]{Question}
\newtheorem{example}[theorem]{Example}

\frenchspacing
%\def\baselinestretch{0.975}

%\emnlpfinalcopy
%\def\emnlppaperid{496}

\title{Word Orders Trade Off Predictability and Memory Load Across Languages}%  Word Order Minimizes  Across Languages
\author{Michael Hahn}
\date{2018}

\begin{document}

\maketitle

\begin{tabular}{cccccccccccccc}
Afrikaans & Amharic & Arabic & Armenian & Basque & Breton & Bulgarian & Buryat \\
\includegraphics[width=0.12\textwidth]{figures/Afrikaans-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Amharic-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Arabic-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Armenian-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Basque-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Breton-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Bulgarian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Buryat-Adap-listener-surprisal-memory.pdf}
\\
Cantonese & Catalan & Chinese & Croatian & Czech & Danish & Dutch & Estonian \\
\includegraphics[width=0.12\textwidth]{figures/Cantonese-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Catalan-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Chinese-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Croatian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Czech-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Danish-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Dutch-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Estonian-listener-surprisal-memory.pdf}
\\
Faroese & Finnish & French & German & Greek & Hebrew & Hindi & Hungarian \\
\includegraphics[width=0.12\textwidth]{figures/Faroese-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Finnish-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/French-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/German-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Greek-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Hebrew-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Hindi-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Hungarian-listener-surprisal-memory.pdf}
\\
Indonesian & Japanese & Kazakh & Kurmanji & Latvian & Naija & North Sami & Norwegian \\
\includegraphics[width=0.12\textwidth]{figures/Indonesian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Japanese-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Kazakh-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Kurmanji-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Latvian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Naija-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/North_Sami-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Norwegian-listener-surprisal-memory.pdf}
\\
Persian & Polish & Romanian & Serbian & Slovak & Slovenian & Spanish & Swedish \\
\includegraphics[width=0.12\textwidth]{figures/Persian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Polish-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Romanian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Serbian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Slovak-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Slovenian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Spanish-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Swedish-listener-surprisal-memory.pdf}
\\
Tamil & Thai & Turkish & Ukrainian & Uyghur & Vietnamese \\
\includegraphics[width=0.12\textwidth]{figures/Tamil-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Thai-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Turkish-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Ukrainian-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Uyghur-Adap-listener-surprisal-memory.pdf} & \includegraphics[width=0.12\textwidth]{figures/Vietnamese-listener-surprisal-memory.pdf}
\\\end{tabular}

\end{document}