### 2019-03-01 results-table-word-level.tex
results

Afrikaans  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Afrikaans-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Afrikaans-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Amharic-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Amharic-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Amharic-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Arabic  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Arabic-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Arabic-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Armenian-Adap*  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Armenian-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Armenian-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.91  &  [0.83, 0.98]  \\ [10.25ex] \hline
Bambara-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Bambara-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Bambara-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Basque  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Basque-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Basque-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Breton-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Breton-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Breton-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Bulgarian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Bulgarian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Bulgarian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Buryat-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Buryat-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Buryat-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Cantonese-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Cantonese-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Cantonese-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.96  &  [0.86, 1.0]  \\ [10.25ex] \hline
Catalan  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Catalan-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Catalan-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Chinese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Chinese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Chinese-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Croatian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Croatian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Croatian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Czech  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Czech-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Czech-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Danish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Danish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Danish-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Dutch  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Dutch-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Dutch-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
English  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/English-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/English-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Erzya-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Erzya-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Erzya-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.99  &  [0.98, 1.0]  \\ [10.25ex] \hline
Estonian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Estonian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Estonian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.83  &  [0.76, 0.9]  \\ [10.25ex] \hline
Faroese-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Faroese-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Faroese-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Finnish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Finnish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Finnish-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
French  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/French-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/French-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
German  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/German-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/German-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Greek  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Greek-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Greek-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hebrew  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hebrew-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hebrew-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hindi  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hindi-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hindi-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hungarian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hungarian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hungarian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.89  &  [0.81, 0.96]  \\ [10.25ex] \hline
Indonesian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Indonesian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Indonesian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Japanese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Japanese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Japanese-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Kazakh-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kazakh-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kazakh-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Kurmanji-Adap*  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kurmanji-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kurmanji-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.92  &  [0.83, 0.99]  \\ [10.25ex] \hline
Latvian*  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Latvian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Latvian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.48  &  [0.4, 0.57]  \\ [10.25ex] \hline
Maltese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Maltese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Maltese-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Naija-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Naija-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Naija-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
North Sami  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/North_Sami-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/North_Sami-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.37  &  [0.3, 0.44]  \\ [10.25ex] \hline
Norwegian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Norwegian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Norwegian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Persian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Persian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Persian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Polish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Polish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Polish-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.07  &  [0.01, 0.15]  \\ [10.25ex] \hline
Portuguese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Portuguese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Portuguese-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Romanian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Romanian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Romanian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Serbian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Serbian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Serbian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Slovak  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovak-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovak-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.07  &  [0.03, 0.12]  \\ [10.25ex] \hline
Slovenian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovenian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovenian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.82  &  [0.74, 0.89]  \\ [10.25ex] \hline
Spanish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Spanish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Spanish-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Swedish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Swedish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Swedish-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Thai-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Thai-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Thai-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Turkish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Turkish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Turkish-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Ukrainian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Ukrainian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Ukrainian-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Urdu  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Urdu-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Urdu-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Uyghur-Adap*  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Uyghur-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Uyghur-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  0.65  &  [0.57, 0.74]  \\ [10.25ex] \hline
Vietnamese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Vietnamese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Vietnamese-listener-surprisal-memory.pdf}}  &  $D_x$  \\ 
  &    &    &  $W_x$  &  1.0  &  [0.98, 1.0]  \\ [10.25ex] \hline