### 2019-03-16 results-table-word-level.tex

Afrikaans  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Afrikaans-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Afrikaans-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Afrikaans-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Afrikaans-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Afrikaans-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1315  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Amharic-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Amharic-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Amharic-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Amharic-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Amharic-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Amharic-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
974  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Arabic  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Arabic-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Arabic-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Arabic-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Arabic-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Arabic-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
21864  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Armenian-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Armenian-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Armenian-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Armenian-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Armenian-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Armenian-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
514  &    &    &    &    &    &  0.92  &  [0.87, 0.97]  \\ [10.25ex] \hline
Bambara-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bambara-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bambara-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bambara-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bambara-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bambara-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
926  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Basque  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Basque-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Basque-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Basque-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Basque-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Basque-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
5396  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Breton-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Breton-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Breton-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Breton-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Breton-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Breton-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
788  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Bulgarian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bulgarian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bulgarian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bulgarian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bulgarian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Bulgarian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
8907  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Buryat-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Buryat-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Buryat-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Buryat-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Buryat-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Buryat-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
808  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Cantonese-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Cantonese-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Cantonese-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Cantonese-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Cantonese-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Cantonese-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
550  &    &    &    &    &    &  0.96  &  [0.86, 1.0]  \\ [10.25ex] \hline
Catalan  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Catalan-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Catalan-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Catalan-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Catalan-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Catalan-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
13123  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Chinese  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Chinese-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Chinese-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Chinese-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Chinese-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Chinese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
3997  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Croatian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Croatian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Croatian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Croatian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Croatian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Croatian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
7689  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Czech  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Czech-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Czech-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Czech-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Czech-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Czech-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
102993  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Danish  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Danish-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Danish-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Danish-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Danish-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Danish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
4383  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Dutch  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Dutch-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Dutch-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Dutch-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Dutch-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Dutch-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
18310  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
English  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/English-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/English-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/English-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/English-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/English-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
17062  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Erzya-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Erzya-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Erzya-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Erzya-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Erzya-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Erzya-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1450  &    &    &    &    &    &  0.99  &  [0.98, 1.0]  \\ [10.25ex] \hline
Estonian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Estonian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Estonian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Estonian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Estonian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Estonian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
6959  &    &    &    &    &    &  0.8  &  [0.72, 0.86]  \\ [10.25ex] \hline
Faroese-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Faroese-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Faroese-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Faroese-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Faroese-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Faroese-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1108  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Finnish  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Finnish-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Finnish-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Finnish-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Finnish-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Finnish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
27198  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
French  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/French-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/French-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/French-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/French-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/French-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
32347  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
German  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/German-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/German-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/German-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/German-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/German-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
13814  &    &    &    &    &    &  1.0  &  [0.91, 1.0]  \\ [10.25ex] \hline
Greek  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Greek-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Greek-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Greek-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Greek-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Greek-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1662  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hebrew  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hebrew-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hebrew-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hebrew-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hebrew-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hebrew-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
5241  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hindi  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hindi-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hindi-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hindi-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hindi-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hindi-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
13304  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hungarian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hungarian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hungarian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hungarian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hungarian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Hungarian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
910  &    &    &    &    &    &  0.87  &  [0.8, 0.93]  \\ [10.25ex] \hline
Indonesian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Indonesian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Indonesian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Indonesian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Indonesian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Indonesian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
4477  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Italian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Italian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Italian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Italian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Italian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Italian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
17427  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Japanese  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Japanese-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Japanese-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Japanese-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Japanese-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Japanese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
7164  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Kazakh-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kazakh-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kazakh-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kazakh-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kazakh-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kazakh-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
947  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Korean  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Korean-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Korean-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Korean-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Korean-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Korean-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
27410  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Kurmanji-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kurmanji-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kurmanji-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kurmanji-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kurmanji-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Kurmanji-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
634  &    &    &    &    &    &  0.93  &  [0.88, 0.98]  \\ [10.25ex] \hline
Latvian*  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Latvian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Latvian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Latvian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Latvian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Latvian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
4124  &    &    &    &    &    &  0.49  &  [0.4, 0.57]  \\ [10.25ex] \hline
Maltese  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Maltese-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Maltese-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Maltese-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Maltese-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Maltese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1123  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Naija-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Naija-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Naija-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Naija-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Naija-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Naija-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
848  &    &    &    &    &    &  1.0  &  [0.99, 1.0]  \\ [10.25ex] \hline
North Sami  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/North_Sami-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/North_Sami-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/North_Sami-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/North_Sami-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/North_Sami-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
2257  &    &    &    &    &    &  0.37  &  [0.3, 0.44]  \\ [10.25ex] \hline
Norwegian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Norwegian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Norwegian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Norwegian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Norwegian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Norwegian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
29870  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Persian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Persian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Persian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Persian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Persian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Persian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
4798  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Polish  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Polish-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Polish-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Polish-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Polish-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Polish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
6100  &    &    &    &    &    &  0.1  &  [0.04, 0.17]  \\ [10.25ex] \hline
Portuguese  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Portuguese-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Portuguese-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Portuguese-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Portuguese-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Portuguese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
17995  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Romanian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Romanian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Romanian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Romanian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Romanian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Romanian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
8664  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Russian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Russian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Russian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Russian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Russian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Russian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
52664  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Serbian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Serbian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Serbian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Serbian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Serbian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Serbian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
2935  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Slovak  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovak-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovak-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovak-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovak-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovak-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
8483  &    &    &    &    &    &  0.07  &  [0.03, 0.12]  \\ [10.25ex] \hline
Slovenian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovenian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovenian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovenian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovenian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Slovenian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
7532  &    &    &    &    &    &  0.82  &  [0.77, 0.88]  \\ [10.25ex] \hline
Spanish  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Spanish-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Spanish-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Spanish-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Spanish-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Spanish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
28492  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Swedish  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Swedish-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Swedish-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Swedish-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Swedish-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Swedish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
7041  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Thai-Adap  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Thai-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Thai-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Thai-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Thai-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Thai-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
900  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Turkish  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Turkish-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Turkish-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Turkish-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Turkish-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Turkish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
3685  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Ukrainian  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Ukrainian-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Ukrainian-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Ukrainian-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Ukrainian-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Ukrainian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
4506  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Urdu  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Urdu-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Urdu-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Urdu-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Urdu-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Urdu-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
4043  &    &    &    &    &    &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Uyghur-Adap*  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Uyghur-Adap-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Uyghur-Adap-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Uyghur-Adap-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Uyghur-Adap-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Uyghur-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1656  &    &    &    &    &    &  0.65  &  [0.57, 0.73]  \\ [10.25ex] \hline
Vietnamese  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Vietnamese-listener-surprisal-memory-by-run_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Vietnamese-listener-surprisal-memory-HIST_byMem_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Vietnamese-listener-surprisal-memory-MEDIANS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Vietnamese-listener-surprisal-memory-MEDIAN_DIFFS_onlyWordForms_boundedVocab.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.1\textwidth]{neural/figures/Vietnamese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab.pdf}}  &    \\ 
1400  &    &    &    &    &    &  1.0  &  [0.98, 1.0]  \\ [10.25ex] \hline