### 2019-08-05 introduction.tex
revising

\mhahn{pasted from the other draft}
Each human language is a system for encoding an infinite variety of thoughts into sentences, i.e. linear strings of words.
A primary goal of the field of linguistics is to characterize and explain the commonalities and differences among these systems, and so to gain insight into the cognitive machinery that underpins this fundamental aspect of human intelligence.
Here we argue that human languages can be modeled as solutions to the problem of maximally efficient communication among agents with particular information-processing constraints active during speaking and listening.
We formalize these constraints using information theory, and show novel large-scale quantitative evidence from 54 languages that information is structured in time so as to minimize memory load during language comprehension.

The idea that the universal properties of human languages are best explained with reference to communicative efficiency has a long history \cite{gabelentz1901sprachwissenschaft,zipf1949human,hockett1960origin,givon1991markedness,hawkins1994performance,hawkins2004efficiency,hawkins2014crosslinguistic,croft2001functional,haspelmath2008parametric,jaeger2011language,gibson2019efficiency}.

These strategies can be seen as sets of rules that differ in the way ...
Languages differ considerably in the rules they apply to order information: English orders the object after the verb, Japanese places it before the verb \note{make more explicit here}.
Explaining the variation in the strategies languages use to  order underlying hierarchical structures into linear strings of words has been one of the foci of linguistic research (\jd{CITE}).
We show that these strategies reflect an underlying optimization principle: human languages are adapted to limitations in human working memory.

%We test the hypothesis that human languages represent different solutions to the problem of efficient computation with constrained working memory.
The suggestion that the structure of human language partly reflects a need for efficient processing under resource limitations has been present in the theoretical and functional linguistics literature for decades~\cite{berwick1984grammatical,hawkins1994performance}, where theories have been developed in terms of specific mechanistic and heuristic models of human language processing.
Here

\mhahn{original intro}
Since the 1950s, it has been a persistent suggestion that human language processing is shaped by a resource bottleneck in short-term memory.
Language is produced and comprehended incrementally in a way that crucially requires both speaker and listener to use an active memory store to keep track of what was previously said.
Since short-term memory of this kind is known to be highly limited in capacity \citep{miller1956magical}, it makes sense for these capacity limits to comprise a major constraint on production and comprehension.
Indeed, a great deal of work in sentence processing has focused on characterizing the effects of memory constraints on language processing \citep{gibson-linguistic-1998,lewis-activation-based-2005,levy2013memory}.

At the same time, the field of functional linguistics has argued that these resource constraints not only affect online language processing, they also shape the form of human language itself.
For example, the Performace--Grammar Correspondence Hypothesis (PGCH) of \citet{hawkins1994performance} holds that forms which are practically easier to produce and comprehend end up becoming part of the grammars of languages, and that this process can explain several of the universal properties of human languages originally documented by \citet{greenberg-universals-1963}.

Here we take up the question of how to characterize short-term memory capacity limitations in language processing for both speakers and listeners, and the question of whether natural language grammars are shaped by these limitations.
Whereas previous theories were based on specific mechanistic models of memory, our theory is purely information-theoretic, meaning that our predictions will hold independently across a wide variety of implementations and architectures.

Our main new concept is the idea of a \emph{memory--surprisal tradeoff}: it is possible for a listener to achieve greater ease of word-by-word comprehension at the cost of investing more computational resources into remembering previous words, and the particular shape of the resulting tradeoff of depends on the word order properties of a language. Analogous results also hold for language production by resource-constrained speakers. We show evidence that the preferred word orders of natural languages are those that enable efficient memory--surprisal tradeoffs.

The remainder of this paper is structured as follows. In Section~\ref{sec:background}, we give a brief review of previous work on the effects of short-term memory on languages and language processing. In Section~\ref{sec:ms-tradeoff}, we describe the memory--surprisal tradeoff and how it results from rate--distortion theory, the theory of optimal information processing under resource constraints \citep{cover2006elements}. In Section~\ref{sec:info-locality}, We prove that word orders enable efficient processing in terms of the memory--surprisal tradeoff when they exhibit \emph{information locality}: whenever utterance elements that predict each other are close to each other. We argue that information locality is a good model of the effects of memory constraints on language processing. In Section~\ref{sec:experiment1}, languages which have previously been shown to be preferred in artificial language experiments are exactly those that enable efficient memory--surprisal tradeoffs \citep{fedzechkina2017human}. In Section~\ref{sec:experiment2}, we show that word orders of natural languages as found in dependency corpora \citep{ud} enable more efficient memory--surprisal tradeoffs than baseline word orders. Section~\ref{sec:conclusion} concludes.