### 2019-08-12 writeup/tables/quantiles_REAL_NGRAMS_0.tex

Afrikaans & Amharic & Arabic & Armenian
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Afrikaans-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Amharic-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Arabic-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Armenian-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Bambara & Basque & Breton & Bulgarian
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Bambara-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Basque-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Breton-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Bulgarian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Buryat & Cantonese & Catalan & Chinese
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Buryat-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Cantonese-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Catalan-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Chinese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Croatian & Czech & Danish & Dutch
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Croatian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Czech-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Danish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Dutch-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
English & Erzya & Estonian & Faroese
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/English-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Erzya-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Estonian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Faroese-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\