### 2019-08-12 writeup/tables/quantiles_REAL_NGRAMS_1.tex

Finnish & French & German & Greek
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Finnish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/French-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/German-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Greek-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Hebrew & Hindi & Hungarian & Indonesian
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Hebrew-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Hindi-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Hungarian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Indonesian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Italian & Japanese & Kazakh & Korean
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Italian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Japanese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Kazakh-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Korean-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Kurmanji & Latvian & Maltese & Naija
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Kurmanji-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Latvian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Maltese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Naija-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
North Sami & Norwegian & Persian & Polish
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/North_Sami-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Norwegian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Persian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Polish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\