### 2019-08-12 writeup/tables/quantiles_REAL_NGRAMS_2.tex

Portuguese & Romanian & Russian & Serbian
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Portuguese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Romanian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Russian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Serbian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Slovak & Slovenian & Spanish & Swedish
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Slovak-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Slovenian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Spanish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Swedish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Thai & Turkish & Ukrainian & Urdu
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Thai-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Turkish-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Ukrainian-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Urdu-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf}
 \\ 
Uyghur & Vietnamese &  & 
 \\ 
\includegraphics[width=0.25\textwidth]{ngrams/figures/Uyghur-Adap-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} & \includegraphics[width=0.25\textwidth]{ngrams/figures/Vietnamese-listener-surprisal-memory-QUANTILES_onlyWordForms_boundedVocab_REAL_NGRAMS.pdf} &  & 
 \\