### 2020-01-14 writeup/si.tex
code

\documentclass[11pt,letterpaper]{article}

\usepackage{showlabels}
\usepackage{fullpage}
\usepackage{pslatex}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{amssymb}

\usepackage{caption}

\usepackage{tikz-dependency}
\usepackage{longtable}

\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\E}[0]{\mathbb{E}}
\newcommand{\Ff}[0]{\mathcal{F}}

\usepackage{multirow}

\newcommand{\soft}[1]{}
\newcommand{\nopreview}[1]{}
\newcommand\comment[1]{{\color{red}#1}}
\newcommand\mhahn[1]{{\color{red}(#1)}}
\newcommand\note[1]{{\color{red}(#1)}}
\newcommand\jd[1]{{\color{red}(#1)}}
\newcommand\rljf[1]{{\color{red}(#1)}}
\newcommand{\key}[1]{\textbf{#1}}

\usepackage{amsthm}

\newcommand{\thetad}[0]{{\theta_d}}
\newcommand{\thetal}[0]{{\theta_{LM}}}

\newcounter{theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{thm}[theorem]{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{question}[theorem]{Question}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}

\frenchspacing
%\def\baselinestretch{0.975}

%\emnlpfinalcopy
%\def\emnlppaperid{496}

\title{Supplementary Information for: Crosslinguistic Word Orders Enable an Efficient Tradeoff between Memory and Surprisal}
\author{Michael Hahn, Judith Degen, Richard Futrell}
\date{2018}

\begin{document}

\maketitle

\begin{center}
\includegraphics[width=0.5\textwidth]{../code/visualize_neural/figures/full-REAL-listener-surprisal-memory-HIST_z_byMem_onlyWordForms_boundedVocab.pdf}
\captionof{figure}{Histogram}\label{fig:hist-real}
\end{center}

\section{Formal Analysis and Proofs}

In this section, we prove the theorem described above.

\subsection{Mathematical Assumptions}

We first make explicit how we formalize language processing for proving the theorem.

\paragraph{Ingredient 1: Language as a Stationary Stochastic Process}
We represent language as a stochastic process of words $\dots w_{-2} w_{-1} w_0 w_{1} w_{2} \dots$, extending indefinitely both into the past and into the future.
The symbols $w_i$ belong to a common set, representing the words of the language.\footnote{Could also be phonemes, sentences, ..., any other kind of unit.}

%We model the sequence as a probabilistic sequence; that is, given a context $w_{<t}$, the next word is distributed according to a distribution $p(w_t|w_{<t})$.

The assumption of infinite length is for mathematical convenience and does not affect the substance of our results:
As we restrict our attention to the processing of individual sentences, which have finite length, we will actually not make use of long-range and infinite contexts.

We make the assumption that this process is \emph{stationary}.
Formally, this means that the conditional distribution $P(w_t|w_{<t})$ does not depend on $t$, it only depends on the actual sequence $w_{<t}$.
Informally, this says that the process has no `internal clock', and that the statistical rules of the language do not change at the timescale we are interested in.
In reality, the statistical rules of language do change: They change as language changes over generations, and they also change between different situations -- e.g., depending on the interlocutor at a given point in time.
Given that we are interested in memory needs in the processing of \emph{individual sentences}, at a timescale of seconds or minutes, stationarity seems to be a reasonable assumption to make.

\begin{figure}
\includegraphics[width=0.45\textwidth]{figures/markov-condition.png}
	\caption{Illustration of (\ref{eq:listener-markov}). As the utterance unfolds, the listener maintains a memory state. After receiving word $w_t$, the listener computes their new memory state $m_t$ based on the previous memory state $m_{t-1}$ and the new word $w_t$.}\label{fig:listener-markov}
\end{figure}

\paragraph{Ingredient 2: Flow of Information}
We now analyze memory from the perspective of the listener, who needs to maintain information about the past to predict the future.
As the speaker's utterance unfolds, the listener maintains a memory state $m_t$.

There are no assumptions about the memory architecture and the nature of its computations.
We only make a basic assumption about the flow of information (Figure~\ref{fig:listener-markov}):
At a given point in time, the listener's memory state $m_t$ is determined by the last word $w_t$, and the prior memory state $m_{t-1}$.
As a consequence, $m_t$ contains no information about the process beyond what is contained in the last word observed $w_{t-1}$ and in the memory state before that word was observed $m_{t-1}$.
This is formalized as a statement about conditional probabilities:
	\begin{equation}\label{eq:listener-markov}
		p(m_1| (w_{t})_{t \in \mathbb{Z}}, m_0)   = p(m_1 | m_0, w_1)
	\end{equation}
This says that $m_1$ contains no information about the utterances beyond what is contained in $m_0$ and $w_1$.	
As a consequence, the listener has no knowledge of the speaker's state beyond the information provided in their prior communication.
This is a simplification, as the listener could obtain information about the speaker from other sources, such as their common environment (weather, ...).
\mhahn{For the study of memory in sentence processing, this seems fair. Discuss this more.}

%
%First, we assume that the listener's internal state cannot depend on the future beyond its dependency on the past.
%Formally: 
%\begin{equation}\label{eq:listener-markov-1}
%m_t \bot w_{>t} | w_{\leq t}
%\end{equation}
%This means that the listener has no access to the speaker's state beyond what the speaker has already uttered.
%
%Second, we assume that $m_t$ contains no information about the past beyond what is contained in $w_{t-1}$ and $m_{t-1}$:
%\begin{equation}\label{eq:listener-markov-2}
%m_t \bot w_{<t} | w_{t-1}, m_{t-1}
%\end{equation}
%This means that any information about the past in $m_t$ has to be contained in $m_{t-1}$ -- formalizing the idea that a listener can only remember aspects of the past by keeping them in memory, and that memories of the past cannot `spontaneously' form later in the future.

%The listener can trade off memory and future surprisal:
%A listener who chooses to store less memory will exerience higher surprisal in the future.
%A listener can achieve minimal surprisal -- that is, the lowest average surprisal that any model could achieve by predicting the future from the past -- if and only if $m_t$ contains all predictive information about the future that is contained in the past.

%We now describe the memory-surprisal tradeoff. will describe this tradeoff, and show that listener memory is linked to locality in a way similar to speaker memory.
%Consider a listener who uses $J$ bits of memory on average.
%What can we say about the listener's surprisal?

\subsection{Proof of the Theorem}

We restate the theorem:

\begin{thm}\label{prop:suboptimal}
	Let $T$ be any positive integer ($T \in \{1, 2, 3, ...\}$), and consider a listener using at most
	\begin{equation}\label{eq:memory}
		\sum_{t=1}^T t I_t
	\end{equation}
bits of memory on average.
Then this listener will incur surprisal at least
	$$H[w_t|w_{<t}] + \sum_{t > T} I_t$$
	on average.
\end{thm}

We formalize a language as a stationary stochastic process $\dots w_{-2} w_{-1} w_0 w_{1} w_{2} \dots$, extending indefinitely both into the past and into the future.
The symbols $w_i$ belong to a common set, representing the words of the language.\footnote{Could also be phonemes, sentences, ..., any other kind of unit.}
We denote the listener's memory state at time $t$, after hearing $w_{<t} = ... w_{t-2} w_{t-1}$ by $m_t$.
As described above, we assume
\begin{equation}\label{eq:listener-markov}
	p(m_{t+1}| (w_{t'})_{t' \in \mathbb{Z}}, m_t)   = p(m_{t+1} | m_t, w_{t})
\end{equation}
that is, $m_{t+1}$ contains no information about the utterances beyond what is contained in $m_t$ and $w_{t}$.
As a consequence, the listener has no knowledge of the speaker's state beyond the information provided in their prior communication.

The average number of bits required to encode this state is $\operatorname{H}[m_t]$, which by assumption is at most $\sum_{t=1}^T t I_t$.
As the listener's predictions are made on the basis of her memory state, her average surprisal is at least $\operatorname{H}[w_t | m_t]$.
The difference between the listener's surprisal and optimal surprisal is thus at least $\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]$.
By the assumption of stationarity, we can, for any positive integer $T$, rewrite this expression as
\begin{equation}\label{eq:byStation}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}] =  \frac{1}{T} \sum_{t'=1}^{T} \left(\operatorname{H}[w_{t'} | m_{t'}] - \operatorname{H}[w_{t'} | w_{<t'}]\right) 
\end{equation}

We first show a lemma:

\begin{lemma}
For any positive integer $t$, the following inequality holds:
\begin{equation}
H[w_t | m_t] \geq H[w_t|w_{1 \dots t-1}, m_1]
\end{equation}
\end{lemma}
\begin{proof}[Proof of the Lemma]
	By Bayes' Theorem
\begin{align*}
	p(w_t|m_0, m_1, w_{0\dots t-1}) &= \frac{p(m_1|m_0, w_{0\dots t})}{p(m_1|m_0, w_{0\dots t-1})} \cdot p(w_t|m_0, w_{0\dots t-1})
\end{align*}
By Equation~\ref{eq:listener-markov}, the quotient on the RHS is equal to $1$, so
\begin{align*}
p(w_t|m_0, m_1, w_{0\dots t-1}) = p(w_t|m_0, w_{0\dots t-1})
\end{align*}
So we have a Markov chain
\begin{equation}
(w_t) \rightarrow (m_0, w_{0 \dots t-1})   \rightarrow   (m_1, w_{1 \dots t-1})
\end{equation}
Thus, by the Data Processing Inequality,
\begin{equation}
H[w_t| w_{1 \dots t-1}, m_{1}] \geq H[w_t|w_{0 \dots t-1}, m_0]
\end{equation}
Finally, iteratively applying this inequality, we get:
\begin{align*}
H[w_t | m_t] \geq H[w_t| w_{t-1}, m_{t-1}] \geq H[w_t| w_{t-2, t-1}, m_{t-2}] \geq ... \geq H[w_t|w_{1 \dots t-1}, m_1]
\end{align*}
\end{proof}
Plugging this inequality into Equation~\ref{eq:byStation} above:
\begin{align*}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]& \geq \frac{1}{T} \sum_{t=1}^T ( \operatorname{H}[w_t|w_{1\dots t-1}, m_1] - \operatorname{H}[w_t | w_{1\dots t-1}, w_{\leq 0}]  )    \\
& = \frac{1}{T} \left(\operatorname{H}[w_{1\dots T} | m_1] - \operatorname{H}[w_{1\dots T} | w_{\leq 0}]\right)  \\
& = \frac{1}{T} \left(I[w_{1\dots T}|w_{\leq 0}] - I[w_{1\dots T}|m_1]\right) 
\end{align*}
The first term $I[w_{1\dots T}|w_{\leq 0}]$ can be rewritten in terms of $I_t$:
\begin{align*}
I[w_{1\dots T}|w_{\leq 0}] &= \sum_{i=1}^T \sum_{j=-1}^{-\infty} I[w_i, w_j | w_{j+1}...w_{i-1}] = \sum_{t=1}^T t I_t + T \sum_{t > T} I_t
\end{align*}
Therefore
\begin{align*}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]& \geq \frac{1}{T} \left(\sum_{t=1}^T t I_t + T \sum_{t > T} I_t - I[w_{1\dots T}|m_1]\right) 
\end{align*}
$I[w_{1\dots T}|m_1]$ is at most $\operatorname{H}[m_1]$, which is at most $\sum_{t=1}^T t I_t$ by assumption. Thus, the expression above is bounded by
\begin{align*}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]& \geq \frac{1}{T} \left(\sum_{t=1}^T t I_t + T \sum_{t > T} I_t - \sum_{t=1}^T t I_t\right) \\
&= \sum_{t > T} I_t
\end{align*}
Rearranging shows that the listener's surprisal is at least $\operatorname{H}[w_t|m_t] \geq \operatorname{H}[w_t | w_{<t}] + \sum_{t > T} I_t$, as claimed.
%\end{proof}

%Justify linear interpolation: The curve is convex, which is shown by `time-sharing': Use one code $\lambda$ fraction of times, and the other code $1-\lambda$ fraction of times.

\subsection{Locality in a model with Memory Retrieval}

Here we show that our information-theoretic analysis is compatible with models placing the main bottleneck in the difficulty of retrieval.
We describe an information-theoretic formalization of a model that has both a limited WM and an unlimited STM.

We consider a model that maintains both a small working memory $m_t$ and an unlimited short-term memory $s_t$.

Predictions are made based on working memory $m_t$, incurring surprisal $H[w_t|m_t]$.
In each time step, there is some amount of communication between $m_t$ and $s_t$, corresponding to retrieval operations.
We model this using a variable $r_t$ representing the information that is retrieved from $s_t$.

This is instantiated by ACT-R.
We can explicitly explain how this covers the McElree ideas and the Lewis and Vasishth ACT-R model.

This model has two bottlenecks:
The working memory capacity, which we model as $H[m_t]$, and the amount of information that is added through retrieval, modeled as $H[r_t|m_t]$.

%Bounding retrieval = bounding the precision and/or content of retrieved items. Explain more how this relates to McElree and ACT-R.

\begin{thm}
	Assume the average working memory $H[m_t]$ is bounded as $H[L_t] \leq \sum_{t=1}^T t I_t$, and the average amount of retrieval is bounded as $H[R_t] \leq \sum_{t=T+1}^S I_t$ (per word).
	Then $H[w_t|m_t] \geq H[w_t|x_{<t}] + \sum_{t>S} I_t$.
\end{thm}

\begin{proof}
The negative surprisal gap is
$\leq \frac{1}{T}(I[X_1\dots X_T|(M_0, R_1, ..., R_T)] - \sum_{t=1}^T t I_t - T \sum_{t>T} I_t) \leq  \frac{1}{T}(T\cdot H[R_t] - T \sum_{t>T} I_t) = H[R_t] - \sum_{t>T} I_t$

\end{proof}

We can also think of this as a multi-objective optimization, aiming to minimize $\lambda_1 WM + \lambda_2 Retrieval$ to achieve a given surprisal level.

If $\frac{\lambda_2}{\lambda_1} \rightarrow \infty$ (retrievals get more expensive), recover previous model.

If $\frac{\lambda_2}{\lambda_1} \rightarrow 0$ (retrievals get cheaper), locality effect gets weaker, and disappears in the limit\footnote{(Of course, even in this limit, there might be additional factors that may still favor locality in a specific implementation of memory -- e.g., in ACT-R, decay and interference are less problematic if there is locality.)}

Objective
$$\min_{T} \lambda_1 \sum_{t=1}^T t I_t + \lambda_2 \sum_{t=T+1}^S I_t$$
has solution $T \approx \frac{\lambda_2}{\lambda_1}$.\footnote{Can do simple proof using the continuous-$T$-version.}

Consequence: As long as retrievals are more expensive than keeping the same amount in WM, locality is predicted.

\section{Corpus Size per Language}

\begin{center}
\begin{longtable}{l|ll||l|llllllllllllll}
	Language & Training & Held-Out & 	Language & Training & Held-Out\\ \hline
\input{tables/corpusSizes.tex}
\end{longtable}
	\captionof{table}{Languages, with the number of training and held-out sentences available.}\label{tab:corpora}
\end{center}

\section{Samples Drawn per Language}

\begin{center}
\begin{longtable}{l|ll||l|llllllllllllll}
	Language & Base. & Real & Language & Base. & Real \\ \hline
\input{tables/samplesNumber.tex}
\end{longtable}
	\captionof{figure}{Samples drawn per language according to the precision-dependent stopping criterion.}\label{tab:samples}
\end{center}

\begin{center}
\begin{longtable}{l|lll||l|lllllllllllllll}
	Language & Mean & Lower & Upper & Language & Mean & Lower & Upper \\ \hline
\input{tables/boot_g_REAL.tex}
\end{longtable}
	\captionof{figure}{Bootstrapped estimates for $G$.}\label{tab:boot-g}
\end{center}

\section{Detailed Results per Language}

\subsection{Median Surprisal per Memory Budget}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_0.tex}
\end{longtable}
	\captionof{figure}{Medians: For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians, dashed lines indicate 95 $\%$ confidence intervals for the population median, dotted lines indicate empirical quantiles ($10\%, 20\%, \dots, 80\%, 90\%$). Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians}
\end{center}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_1.tex}
\end{longtable}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_2.tex}
\end{longtable}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_3.tex}
\end{longtable}
	\captionof{figure}{Medians (cont.)}
\end{center}

\subsection{Surprisal at Maximum Memory}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_0.tex}
\end{tabular}
	\captionof{figure}{Histograms: Surprisal, at maximum memory.}\label{tab:slice-hists-real}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_1.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_2.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_3.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

%\begin{table}
%\begin{tabular}{cccccccccccccccccc}
%\input{tables/quantiles_REAL_0.tex}
%\end{tabular}
%	\captionof{figure}{Quantiles: At a given memory budget, what percentage of the baselines results in higher listener surprisal than the real language? Solid curves represent sample means, dashed lines represent 95 \% confidence bounds; dotted lines represent 99.9 \% confidence bounds. At five evenly spaced memory levels, we provide a p-value for the null hypothesis that the actual population mean is $0.5$ or less. Confidence bounds and p-values are obtained using an exact nonparametric method (see text).}\label{tab:quantiles}
%\end{center}
%
%\begin{table}
%\begin{tabular}{cccccccccccccccccc}
%\input{tables/quantiles_REAL_1.tex}
%\end{tabular}
%	\captionof{figure}{Quantiles (part 2)}
%\end{center}
%
%\begin{table}
%\begin{tabular}{cccccccccccccccccc}
%\input{tables/quantiles_REAL_2.tex}
%\end{tabular}
%	\captionof{figure}{Quantiles (part 3)}
%\end{center}
%
%
%

\subsection{Samples Drawn (Experiment 3)}

\begin{center}
\begin{tabular}{l|ll||l|llllllllllllll}
	Language & Base. & MLE & Language & Base. & MLE \\ \hline
\input{tables/samplesNumber_ground.tex}
\end{tabular}
	\captionof{figure}{Experiment 3: Samples drawn per language according to the precision-dependent stopping criterion.}\label{tab:samples}
\end{center}

\subsection{Medians (Experiment 3)}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_0.tex}
\end{tabular}
	\captionof{figure}{Experiment 3. Medians: For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians, dashed lines indicate 95 $\%$ confidence intervals for the population median. Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_1.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_2.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_3.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_0.tex}
\end{tabular}
	\captionof{figure}{Median Differences between Real and Baseline: For each memory budget, we provide the difference in median surprisal between real languages and random baselines; for real orders (blue) and maximum likelihood grammars (red). Lower values indicate lower surprisal compared to baselines. Solid lines indicate sample means. Dashed lines indicate 95 $\%$ confidence intervals.}\label{tab:median_diffs}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_1.tex}
\end{tabular}
	\captionof{figure}{Median Differences (Part 2)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_2.tex}
\end{tabular}
	\captionof{figure}{Median Differences (Part 3)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_3.tex}
\end{tabular}
	\captionof{figure}{Median Differences (Part 4)}
\end{center}

\begin{center}
\begin{tabular}{cccccccccccccccccc}
\input{tables/quantiles_noAssumption_0.tex}
\end{tabular}
	\captionof{figure}{Quantiles: At a given memory budget, what percentage of the baselines results in higher listener surprisal than the real language? Solid curves represent sample means, dashed lines represent 95 \% confidence bounds; dotted lines represent 99.9 \% confidence bounds. At five evenly spaced memory levels, we provide a p-value for the null hypothesis that the actual population mean is $0.5$ or less. Confidence bounds and p-values are obtained using an exact nonparametric method (see text).}\label{tab:quantiles}
\end{center}

\begin{center}
\begin{tabular}{cccccccccccccccccc}
\input{tables/quantiles_noAssumption_1.tex}
\end{tabular}
	\captionof{figure}{Quantiles (part 2)}
\end{center}

\begin{center}
\begin{tabular}{cccccccccccccccccc}
\input{tables/quantiles_noAssumption_2.tex}
\end{tabular}
	\captionof{figure}{Quantiles (part 3)}
\end{center}

\section{Details for Neural Network Models}

\section{N-Gram Models}

\input{control-ngrams.tex}

%-- English, Korean, Russian
%-- UD$\_$Polish-LFG (released in 2.2, not included in original experiment) (13,744 sentences)
%-- character-level Russian
%\section{Character-Level Modeling}
%\section{Non-UD Dependency Treebanks}
%- other treebanks
%-- spoken Japanese (T{\"u}ba-J/S)
%-- another Vietnamese dependency treebank \citep{nguyen-bktreebank:-2017} (5,639 sentences)
%-- another Chinese dependency treebank LDC2012T05
%Due to the sizes of these treebanks, can also do experiment with full word forms.
%
%
%\section{Constituency Treebank}
%
%-- Penn treebank \citep{marcus-building-1993}
%
%-- spoken English (T{\"u}ba-E/S)
%
%-- spoken German (T{\"u}ba-D/S)
%
%-- Chinese treebank \citep{xue-chinese-2013}

\bibliographystyle{apalike}
\bibliography{literature}

%\appendix

\end{document}