### 2020-02-05 writeup/si.tex

\documentclass[11pt,letterpaper]{article}

\usepackage{showlabels}
\usepackage{fullpage}
\usepackage{pslatex}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{amssymb}

\usepackage{caption}

\usepackage{tikz-dependency}
\usepackage{longtable}

\newcommand{\R}[0]{\mathbb{R}}
\newcommand{\E}[0]{\mathbb{E}}
\newcommand{\Ff}[0]{\mathcal{F}}

\usepackage{multirow}

\newcommand{\soft}[1]{}
\newcommand{\nopreview}[1]{}
\newcommand\comment[1]{{\color{red}#1}}
\newcommand\mhahn[1]{{\color{red}(#1)}}
\newcommand\note[1]{{\color{red}(#1)}}
\newcommand\jd[1]{{\color{red}(#1)}}
\newcommand\rljf[1]{{\color{red}(#1)}}
\newcommand{\key}[1]{\textbf{#1}}

\usepackage{amsthm}

\newcommand{\thetad}[0]{{\theta_d}}
\newcommand{\thetal}[0]{{\theta_{LM}}}

\newcounter{theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{thm}[theorem]{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{question}[theorem]{Question}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}

\frenchspacing
%\def\baselinestretch{0.975}

%\emnlpfinalcopy
%\def\emnlppaperid{496}

\title{Supplementary Information for: Crosslinguistic Word Orders Enable an Efficient Tradeoff between Memory and Surprisal}
\author{Michael Hahn, Judith Degen, Richard Futrell}
\date{2018}

\begin{document}

\maketitle

\begin{center}
\includegraphics[width=0.5\textwidth]{../code/analysis/visualize_neural/figures/full-REAL-listener-surprisal-memory-HIST_z_byMem_onlyWordForms_boundedVocab.pdf}
\captionof{figure}{Histogram}\label{fig:hist-real}
\end{center}

\section{Formal Analysis and Proofs}

In this section, we prove the theorem described above.

\subsection{Mathematical Assumptions}

We first make explicit how we formalize language processing for proving the theorem.

\paragraph{Ingredient 1: Language as a Stationary Stochastic Process}
We represent language as a stochastic process of words $\dots w_{-2} w_{-1} w_0 w_{1} w_{2} \dots$, extending indefinitely both into the past and into the future.
The symbols $w_i$ belong to a common set, representing the words of the language.\footnote{Could also be phonemes, sentences, ..., any other kind of unit.}

%We model the sequence as a probabilistic sequence; that is, given a context $w_{<t}$, the next word is distributed according to a distribution $p(w_t|w_{<t})$.

The assumption of infinite length is for mathematical convenience and does not affect the substance of our results:
As we restrict our attention to the processing of individual sentences, which have finite length, we will actually not make use of long-range and infinite contexts.

We make the assumption that this process is \emph{stationary}.
Formally, this means that the conditional distribution $P(w_t|w_{<t})$ does not depend on $t$, it only depends on the actual sequence $w_{<t}$.
Informally, this says that the process has no `internal clock', and that the statistical rules of the language do not change at the timescale we are interested in.
In reality, the statistical rules of language do change: They change as language changes over generations, and they also change between different situations -- e.g., depending on the interlocutor at a given point in time.
Given that we are interested in memory needs in the processing of \emph{individual sentences}, at a timescale of seconds or minutes, stationarity seems to be a reasonable assumption to make.

\begin{figure}
\includegraphics[width=0.45\textwidth]{figures/markov-condition.png}
	\caption{Illustration of (\ref{eq:listener-markov}). As the utterance unfolds, the listener maintains a memory state. After receiving word $w_t$, the listener computes their new memory state $m_t$ based on the previous memory state $m_{t-1}$ and the new word $w_t$.}\label{fig:listener-markov}
\end{figure}

\paragraph{Ingredient 2: Flow of Information}
We now analyze memory from the perspective of the listener, who needs to maintain information about the past to predict the future.
As the speaker's utterance unfolds, the listener maintains a memory state $m_t$.

There are no assumptions about the memory architecture and the nature of its computations.
We only make a basic assumption about the flow of information (Figure~\ref{fig:listener-markov}):
At a given point in time, the listener's memory state $m_t$ is determined by the last word $w_t$, and the prior memory state $m_{t-1}$.
As a consequence, $m_t$ contains no information about the process beyond what is contained in the last word observed $w_{t-1}$ and in the memory state before that word was observed $m_{t-1}$.
This is formalized as a statement about conditional probabilities:
	\begin{equation}\label{eq:listener-markov}
		p(m_1| (w_{t})_{t \in \mathbb{Z}}, m_0)   = p(m_1 | m_0, w_1)
	\end{equation}
This says that $m_1$ contains no information about the utterances beyond what is contained in $m_0$ and $w_1$.	
As a consequence, the listener has no knowledge of the speaker's state beyond the information provided in their prior communication.
This is a simplification, as the listener could obtain information about the speaker from other sources, such as their common environment (weather, ...).
\mhahn{For the study of memory in sentence processing, this seems fair. Discuss this more.}

%
%First, we assume that the listener's internal state cannot depend on the future beyond its dependency on the past.
%Formally: 
%\begin{equation}\label{eq:listener-markov-1}
%m_t \bot w_{>t} | w_{\leq t}
%\end{equation}
%This means that the listener has no access to the speaker's state beyond what the speaker has already uttered.
%
%Second, we assume that $m_t$ contains no information about the past beyond what is contained in $w_{t-1}$ and $m_{t-1}$:
%\begin{equation}\label{eq:listener-markov-2}
%m_t \bot w_{<t} | w_{t-1}, m_{t-1}
%\end{equation}
%This means that any information about the past in $m_t$ has to be contained in $m_{t-1}$ -- formalizing the idea that a listener can only remember aspects of the past by keeping them in memory, and that memories of the past cannot `spontaneously' form later in the future.

%The listener can trade off memory and future surprisal:
%A listener who chooses to store less memory will exerience higher surprisal in the future.
%A listener can achieve minimal surprisal -- that is, the lowest average surprisal that any model could achieve by predicting the future from the past -- if and only if $m_t$ contains all predictive information about the future that is contained in the past.

%We now describe the memory-surprisal tradeoff. will describe this tradeoff, and show that listener memory is linked to locality in a way similar to speaker memory.
%Consider a listener who uses $J$ bits of memory on average.
%What can we say about the listener's surprisal?

\subsection{Proof of the Theorem}

We restate the theorem:

\begin{thm}\label{prop:suboptimal}
	Let $T$ be any positive integer ($T \in \{1, 2, 3, ...\}$), and consider a listener using at most
	\begin{equation}\label{eq:memory}
		\sum_{t=1}^T t I_t
	\end{equation}
bits of memory on average.
Then this listener will incur surprisal at least
	$$H[w_t|w_{<t}] + \sum_{t > T} I_t$$
	on average.
\end{thm}

We formalize a language as a stationary stochastic process $\dots w_{-2} w_{-1} w_0 w_{1} w_{2} \dots$, extending indefinitely both into the past and into the future.
The symbols $w_i$ belong to a common set, representing the words of the language.\footnote{Could also be phonemes, sentences, ..., any other kind of unit.}
We denote the listener's memory state at time $t$, after hearing $w_{<t} = ... w_{t-2} w_{t-1}$ by $m_t$.
As described above, we assume
\begin{equation}
	m_t = M(m_{t-1}, w_{t-1})
\end{equation}
\footnote{Alternatively we could admit nondeterministic memory encodings, and require
\begin{equation}\label{eq:listener-markov}
	p(m_{t+1}| (w_{t'})_{t' \in \mathbb{Z}}, m_t)   = p(m_{t+1} | m_t, w_{t})
\end{equation}
that is, $m_{t+1}$ contains no information about the utterances beyond what is contained in $m_t$ and $w_{t}$.}
As a consequence, the listener has no knowledge of the speaker's state beyond the information provided in their prior communication.

The average number of bits required to encode this state is $\operatorname{H}[m_t]$, which by assumption is at most $\sum_{t=1}^T t I_t$.
As the listener's predictions are made on the basis of her memory state, her average surprisal is at least $\operatorname{H}[w_t | m_t]$.
The difference between the listener's surprisal and optimal surprisal is thus at least $\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]$.
By the assumption of stationarity, we can, for any positive integer $T$, rewrite this expression as
\begin{equation}\label{eq:byStation}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}] =  \frac{1}{T} \sum_{t'=1}^{T} \left(\operatorname{H}[w_{t'} | m_{t'}] - \operatorname{H}[w_{t'} | w_{<t'}]\right) 
\end{equation}

%NOTE the proof could be made simpler by taking $m_t$ to be a deterministic function of $x_t$, $m_{t-1}$, rather than talking about conditional independence
%
%We first show a lemma:
%
%\begin{lemma}
%For any positive integer $t$, the following inequality holds:
%\begin{equation}
%H[w_t | m_t] \geq H[w_t|w_{1 \dots t-1}, m_1]
%\end{equation}
%\end{lemma}
%
%\begin{proof}[Proof of the Lemma]
Because $m_t$ is determined by $(w_{1 \dots t-1}, m_1)$:
\begin{equation}
	m_t = M(m_{t-1}, w_{t-1}) = M(M(m_{t-2}, w_{t-2}), w_{t-1}) = M(M(M(m_{t-3}, w_{t-3}), w_{t-2}), w_{t-1}) = \dots
\end{equation}
the Data Processing inequality entails the following inequality for every positive integer $t$:
\begin{equation}
	H[w_t | m_t] \geq H[w_t|w_{1 \dots t-1}, m_1]
\end{equation}

%\end{proof}

%\begin{proof}[Proof of the Lemma]
%	By Bayes' Theorem
%\begin{align*}
%	p(w_t|m_0, m_1, w_{0\dots t-1}) &= \frac{p(m_1|m_0, w_{0\dots t})}{p(m_1|m_0, w_{0\dots t-1})} \cdot p(w_t|m_0, w_{0\dots t-1})
%\end{align*}
%By Equation~\ref{eq:listener-markov}, the quotient on the RHS is equal to $1$, so
%\begin{align*}
%p(w_t|m_0, m_1, w_{0\dots t-1}) = p(w_t|m_0, w_{0\dots t-1})
%\end{align*}
%So we have a Markov chain
%\begin{equation}
%(w_t) \rightarrow (m_0, w_{0 \dots t-1})   \rightarrow   (m_1, w_{1 \dots t-1})
%\end{equation}
%Thus, by the Data Processing Inequality,
%\begin{equation}
%H[w_t| w_{1 \dots t-1}, m_{1}] \geq H[w_t|w_{0 \dots t-1}, m_0]
%\end{equation}
%Finally, iteratively applying this inequality, we get:
%\begin{align*}
%H[w_t | m_t] \geq H[w_t| w_{t-1}, m_{t-1}] \geq H[w_t| w_{t-2, t-1}, m_{t-2}] \geq ... \geq H[w_t|w_{1 \dots t-1}, m_1]
%\end{align*}
%\end{proof}
Plugging this inequality into Equation~\ref{eq:byStation} above:
\begin{align*}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]& \geq \frac{1}{T} \sum_{t=1}^T ( \operatorname{H}[w_t|w_{1\dots t-1}, m_1] - \operatorname{H}[w_t | w_{1\dots t-1}, w_{\leq 0}]  )    \\
& = \frac{1}{T} \left(\operatorname{H}[w_{1\dots T} | m_1] - \operatorname{H}[w_{1\dots T} | w_{\leq 0}]\right)  \\
& = \frac{1}{T} \left(I[w_{1\dots T}, w_{\leq 0}] - I[w_{1\dots T}, m_1]\right) 
\end{align*}
The first term $I[w_{1\dots T}, w_{\leq 0}]$ can be rewritten in terms of $I_t$:
\begin{align*}
I[w_{1\dots T}, w_{\leq 0}] &= \sum_{i=1}^T \sum_{j=-1}^{-\infty} I[w_i, w_j | w_{j+1}...w_{i-1}] = \sum_{t=1}^T t I_t + T \sum_{t > T} I_t
\end{align*}
Therefore
\begin{align*}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]& \geq \frac{1}{T} \left(\sum_{t=1}^T t I_t + T \sum_{t > T} I_t - I[w_{1\dots T}, m_1]\right) 
\end{align*}
$I[w_{1\dots T}|m_1]$ is at most $\operatorname{H}[m_1]$, which is at most $\sum_{t=1}^T t I_t$ by assumption. Thus, the expression above is bounded by
\begin{align*}
\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]& \geq \frac{1}{T} \left(\sum_{t=1}^T t I_t + T \sum_{t > T} I_t - \sum_{t=1}^T t I_t\right) \\
&= \sum_{t > T} I_t
\end{align*}
Rearranging shows that the listener's surprisal is at least $\operatorname{H}[w_t|m_t] \geq \operatorname{H}[w_t | w_{<t}] + \sum_{t > T} I_t$, as claimed.
%\end{proof}

%Justify linear interpolation: The curve is convex, which is shown by `time-sharing': Use one code $\lambda$ fraction of times, and the other code $1-\lambda$ fraction of times.

\subsection{Locality in a model with Memory Retrieval}

Here we show that our information-theoretic analysis is compatible with models placing the main bottleneck in the difficulty of retrieval.
We describe an information-theoretic formalization of a model that has both a limited WM and an unlimited STM.

We consider a model that maintains both a small working memory $m_t$ and an unlimited short-term memory $s_t$.

Predictions are made based on working memory $m_t$, incurring surprisal $H[w_t|m_t]$.
In each time step, there is some amount of communication between $m_t$ and $s_t$, corresponding to retrieval operations.
We model this using a variable $r_t$ representing the information that is retrieved from $s_t$.

This is instantiated by ACT-R.
We can explicitly explain how this covers the McElree ideas and the Lewis and Vasishth ACT-R model.

This model has two bottlenecks:
The working memory capacity, which we model as $H[m_t]$, and the amount of information that is added through retrieval, modeled as $H[r_t|m_t]$.

Bounding retrieval = bounding the precision and/or content of retrieved items. Explain more how this relates to McElree and ACT-R.

In ACT-R, each retrieval operation takes time. If each chunk has a bounded amount of information, then this corresponds to a bottleneck in $H[r_t|m_t]$

Formally:

$m_t = f(x_t, m_{t-1}, r_t)$

$r_t = g(m_{t-1}, s_{t-1})$

$s_t = h(m_{t-1}, s_{t-1})$

\begin{thm}
Assume the average working memory $H[m_t]$ is bounded as $H[L_t] \leq \sum_{t=1}^T t I_t$, and the average amount of retrieval is bounded as $H[R_t] \leq \sum_{t=T+1}^S I_t$ (per word).
Then $H[w_t|m_t] \geq H[w_t|x_{<t}] + \sum_{t>S} I_t$.
\end{thm}

\begin{proof}
	As before using the Data Processing inequality:
\begin{align*}
	-(\operatorname{H}[w_t | m_t] - \operatorname{H}[w_t | w_{<t}]) \leq & \frac{1}{T}\left(I[X_1\dots X_T, (M_0, R_1, ..., R_T)] - I[X_1\dots X_T, X_{<1}]\right) \\
= & \frac{1}{T}\left(I[X_1\dots X_T, (M_0, R_1, ..., R_T)] - \sum_{t=1}^T t I_t - T \sum_{t>T} I_t\right) \\
	= & \frac{1}{T}\left(I[X_{1\dots T}, M_0] + I[X_{1\dots T}, R_1|M_0] + \dots + I[X_{1\dots T}, R_T|M_0, R_{1 \dots T-1}] - \sum_{t=1}^T t I_t - T \sum_{t>T} I_t\right) \\
	\leq & \frac{1}{T}\left(H[M_0] + H[R_1] + \dots + H[R_{T-1}] - \sum_{t=1}^T t I_t - T \sum_{t>T} I_t\right) \\
	\leq &  \frac{1}{T}(T\cdot H[R_t] - T \sum_{t>T} I_t) \\
	= & H[R_t] - \sum_{t>T} I_t
\end{align*}

\end{proof}

We can also think of this as a multi-objective optimization, aiming to minimize $\lambda_1 WM + \lambda_2 Retrieval$ to achieve a given surprisal level.

If $\frac{\lambda_2}{\lambda_1} \rightarrow \infty$ (retrievals get more expensive), recover previous model.

If $\frac{\lambda_2}{\lambda_1} \rightarrow 0$ (retrievals get cheaper), locality effect gets weaker, and disappears in the limit\footnote{(Of course, even in this limit, there might be additional factors that may still favor locality in a specific implementation of memory -- e.g., in ACT-R, decay and interference are less problematic if there is locality.)}

Objective
$$\min_{T} \lambda_1 \sum_{t=1}^T t I_t + \lambda_2 \sum_{t=T+1}^S I_t$$
has solution $T \approx \frac{\lambda_2}{\lambda_1}$.\footnote{Can do simple proof using the continuous-$T$-version.}

Consequence: As long as retrievals are more expensive than keeping the same amount in WM, locality is predicted.

\subsection{Results for Language Production}

Here we show results linking memory and locality in production.

First, we consider a setting in which a speaker produces sentences with bounded memory, and analyze the deviation of the produced distribution from the actual distribution of the language.

The first setting does not account for the fact that language is produced aiming for some communicative goal.
We therefore now assume that the speaker has a communicative goal $G$ in mind.
This goal $G$ stays constant during production process for a sentence, and we count how much memory is needed in addition to the goal $G$.
We assume that there is a distribution of sentences expressing goals $G$:
\begin{equation}
	P(sentence|G)
\end{equation}
and assume that the speaker aims to match this distribution
\begin{equation}
\mathbb{E}_G[D_{KL}((language|G)||(produced|G))]
\end{equation}
We can analyze this model by adding conditioning w.r.t. $G$ throughout the analysis of the previous case.
Specifically, we need $I_t^G := I[X_t, X_0|X_1, \dots, X_{t-1}, G]$.

Take $I_t$ conditioned on $G$: only count statistical dependencies to the degree that they are not redundant with the goal

\section{Corpus Size per Language}

\begin{center}
\begin{longtable}{l|ll||l|llllllllllllll}
	Language & Training & Held-Out & 	Language & Training & Held-Out\\ \hline
\input{tables/corpusSizes.tex}
\end{longtable}
	\captionof{table}{Languages, with the number of training and held-out sentences available.}\label{tab:corpora}
\end{center}

\section{Samples Drawn per Language}

\begin{center}
\begin{longtable}{l|ll||l|llllllllllllll}
	Language & Base. & Real & Language & Base. & Real \\ \hline
\input{tables/samplesNumber.tex}
\end{longtable}
	\captionof{figure}{Samples drawn per language according to the precision-dependent stopping criterion.}\label{tab:samples}
\end{center}

\begin{center}
\begin{longtable}{l|lll||l|lllllllllllllll}
	Language & Mean & Lower & Upper & Language & Mean & Lower & Upper \\ \hline
\input{tables/boot_g_REAL.tex}
\end{longtable}
	\captionof{figure}{Bootstrapped estimates for $G$.}\label{tab:boot-g}
\end{center}

\section{Detailed Results per Language}

\subsection{Median Surprisal per Memory Budget}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_0.tex}
\end{longtable}
	\captionof{figure}{Medians: For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians, dashed lines indicate 95 $\%$ confidence intervals for the population median, dotted lines indicate empirical quantiles ($10\%, 20\%, \dots, 80\%, 90\%$). Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians}
\end{center}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_1.tex}
\end{longtable}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_2.tex}
\end{longtable}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_REAL_3.tex}
\end{longtable}
	\captionof{figure}{Medians (cont.)}
\end{center}

\subsection{Surprisal at Maximum Memory}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_0.tex}
\end{tabular}
	\captionof{figure}{Histograms: Surprisal, at maximum memory.}\label{tab:slice-hists-real}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_1.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_2.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/slice-hists_REAL_3.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

%\begin{table}
%\begin{tabular}{cccccccccccccccccc}
%\input{tables/quantiles_REAL_0.tex}
%\end{tabular}
%	\captionof{figure}{Quantiles: At a given memory budget, what percentage of the baselines results in higher listener surprisal than the real language? Solid curves represent sample means, dashed lines represent 95 \% confidence bounds; dotted lines represent 99.9 \% confidence bounds. At five evenly spaced memory levels, we provide a p-value for the null hypothesis that the actual population mean is $0.5$ or less. Confidence bounds and p-values are obtained using an exact nonparametric method (see text).}\label{tab:quantiles}
%\end{center}
%
%\begin{table}
%\begin{tabular}{cccccccccccccccccc}
%\input{tables/quantiles_REAL_1.tex}
%\end{tabular}
%	\captionof{figure}{Quantiles (part 2)}
%\end{center}
%
%\begin{table}
%\begin{tabular}{cccccccccccccccccc}
%\input{tables/quantiles_REAL_2.tex}
%\end{tabular}
%	\captionof{figure}{Quantiles (part 3)}
%\end{center}
%
%
%

\subsection{Samples Drawn (Experiment 3)}

\begin{center}
\begin{tabular}{l|ll||l|llllllllllllll}
	Language & Base. & MLE & Language & Base. & MLE \\ \hline
\input{tables/samplesNumber_ground.tex}
\end{tabular}
	\captionof{figure}{Experiment 3: Samples drawn per language according to the precision-dependent stopping criterion.}\label{tab:samples}
\end{center}

\subsection{Medians (Experiment 3)}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_0.tex}
\end{tabular}
	\captionof{figure}{Experiment 3. Medians: For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians, dashed lines indicate 95 $\%$ confidence intervals for the population median. Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_1.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_2.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccclll}
\input{tables/medians_3.tex}
\end{tabular}
	\captionof{figure}{Medians (cont.)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_0.tex}
\end{tabular}
	\captionof{figure}{Median Differences between Real and Baseline: For each memory budget, we provide the difference in median surprisal between real languages and random baselines; for real orders (blue) and maximum likelihood grammars (red). Lower values indicate lower surprisal compared to baselines. Solid lines indicate sample means. Dashed lines indicate 95 $\%$ confidence intervals.}\label{tab:median_diffs}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_1.tex}
\end{tabular}
	\captionof{figure}{Median Differences (Part 2)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_2.tex}
\end{tabular}
	\captionof{figure}{Median Differences (Part 3)}
\end{center}

\begin{center}
\begin{tabular}{ccccccccccccccccll}
\input{tables/medianDiff_3.tex}
\end{tabular}
	\captionof{figure}{Median Differences (Part 4)}
\end{center}

\begin{center}
\begin{tabular}{cccccccccccccccccc}
\input{tables/quantiles_noAssumption_0.tex}
\end{tabular}
	\captionof{figure}{Quantiles: At a given memory budget, what percentage of the baselines results in higher listener surprisal than the real language? Solid curves represent sample means, dashed lines represent 95 \% confidence bounds; dotted lines represent 99.9 \% confidence bounds. At five evenly spaced memory levels, we provide a p-value for the null hypothesis that the actual population mean is $0.5$ or less. Confidence bounds and p-values are obtained using an exact nonparametric method (see text).}\label{tab:quantiles}
\end{center}

\begin{center}
\begin{tabular}{cccccccccccccccccc}
\input{tables/quantiles_noAssumption_1.tex}
\end{tabular}
	\captionof{figure}{Quantiles (part 2)}
\end{center}

\begin{center}
\begin{tabular}{cccccccccccccccccc}
\input{tables/quantiles_noAssumption_2.tex}
\end{tabular}
	\captionof{figure}{Quantiles (part 3)}
\end{center}

\section{Details for Neural Network Models}

\section{N-Gram Models}

\input{control-ngrams.tex}

%-- English, Korean, Russian
%-- UD$\_$Polish-LFG (released in 2.2, not included in original experiment) (13,744 sentences)
%-- character-level Russian
%\section{Character-Level Modeling}
%\section{Non-UD Dependency Treebanks}
%- other treebanks
%-- spoken Japanese (T{\"u}ba-J/S)
%-- another Vietnamese dependency treebank \citep{nguyen-bktreebank:-2017} (5,639 sentences)
%-- another Chinese dependency treebank LDC2012T05
%Due to the sizes of these treebanks, can also do experiment with full word forms.
%
%
%\section{Constituency Treebank}
%
%-- Penn treebank \citep{marcus-building-1993}
%
%-- spoken English (T{\"u}ba-E/S)
%
%-- spoken German (T{\"u}ba-D/S)
%
%-- Chinese treebank \citep{xue-chinese-2013}

\bibliographystyle{apalike}
\bibliography{literature}

%\appendix

\end{document}