### 2020-02-07 writeup/control-ngrams.tex
code

\subsection{Method}
We use a version of Kneser-Ney Smoothing.
For a sequence $w_1\dots w_k$, let $N(w_{1\dots k})$ be the number of times $w_{1\dots k}$ occurs in the training set.
The unigram probabilities are estimated as
\begin{equation}
	p_1(w_t) :=   \frac{N(w_t) + \delta}{|Train| + |V| \cdot \delta}
\end{equation}
where $\delta \in \mathbb{R}_+$ is a hyperparameter.
Here $|Train|$ is the number of tokens in the training set, $|V|$ is the number of types occurring in train or held-out data.
Higher-order probabilities $p_t(w_t|w_{0 \dots t-1})$ are estimated recursively as follows.
Let $\gamma > 0$ be a hyperparameter.
If $N(w_{0 \dots t-1}) < \gamma$, set
\begin{equation}
	p_t(w_t|w_{0 \dots t-1}) := p_{t-1}(w_t|w_{1\dots t-1})
\end{equation}
Otherwise, we interpolate between $t$-th order and lower-order estimates:
\begin{equation}
	p_t(w_t|w_{0 \dots t-1}) :=  \frac{\operatorname{max}(N(w_{0\dots t}) - \alpha, 0.0) + \alpha \cdot \#\{w : N(w_{0 \dots t-1}w) > 0\} \cdot p_{t-1}(w_t|w_{1\dots t-1})}{N(w_{0\dots t-1})}
\end{equation}
where $\alpha \in [0,1]$ is also a hyperparameter.
(CITE) show that this definition results in a well-defined probability distribution, i.e., $\sum_{w \in V} p_t(w|w_{0 \dots t-1}) = 1$.

%Note that this definition guarantees $\sum_{w \in V} p_t(w|w_{0 \dots t-1}) = 1$, because
%
%$\sum_{w \in V} \operatorname{max}(N(w_{0\dots t-1}w) - \alpha, 0.0) + \alpha \cdot \#\{w' : N(w_{0 \dots t-1}w') > 0\} \cdot p_{t-1}(w|w_{1\dots t-1})$
%
%$\sum_{w \in V : N(w_{0\dots t-1} w) > 0} (N(w_{0\dots t}) - \alpha) + \alpha \cdot \sum_{w \in V} \#\{w' : N(w_{0 \dots t-1}w') > 0\} \cdot p_{t-1}(w|w_{1\dots t-1})$
%
%$\sum_{w \in V : N(w_{0\dots t-1} w) > 0} (N(w_{0\dots t}) - \alpha) + \alpha \cdot \#\{w' : N(w_{0 \dots t-1}w') > 0\}$
%
%$\sum_{w \in V} : N(w_{0\dots t-1} w) = N(w_{0\dots t-1})$

Hyperparameters $\alpha, \gamma, \delta$ are tuned with the same strategy as for the neural network models.

\subsection{Results}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_ngrams_0.tex}
\input{tables/medians_ngrams_1.tex}
\input{tables/medians_ngrams_2.tex}
\input{tables/medians_ngrams_3.tex}
\end{longtable}
	\captionof{figure}{Medians (estimated using n-gram models): For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians for ngrams, dashed lines indicate 95 \% confidence intervals for the population median. Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians_ngrams}
\end{center}

% ../../writeup/tables/quantiles_REAL_NGRAMS_0.tex

\begin{center}
\begin{longtable}{cccccccccccccccccc}
\input{tables/quantiles_REAL_NGRAMS_0.tex}
\input{tables/quantiles_REAL_NGRAMS_1.tex}
\input{tables/quantiles_REAL_NGRAMS_2.tex}
\end{longtable}
	\captionof{figure}{Quantiles: At a given memory budget, what percentage of the baselines results in higher listener surprisal than the real language? Solid curves represent sample means, dashed lines represent 95 \% confidence bounds; dotted lines represent 99.9 \% confidence bounds. At five evenly spaced memory levels, we provide a p-value for the null hypothesis that the actual population mean is $0.5$ or less. Confidence bounds and p-values are obtained using an exact nonparametric method (see text).}\label{tab:quantiles}
\end{center}

%bounds.append(["alpha", float, 0.95, 1.0]) # + [x/20.0 for x in range(15, 21)])
%bounds.append(["gamma", int, 1, 2, 3, 4, 5, 8, 10, 15, 20, 25, 30]) # , 200, 300
%bounds.append(["delta", int, 0.1, 0.2, 0.5, 1.0 , 2.0, 3.0, 4.0, 5.0, 8.0, 10.0]) #, 1024]) #, 1024]) # 64, 128,
%bounds.append(["cutoff", int, 2,3,4,5,6,7,8,9,10]) #,7,8,9,10]) #, 1024]) #, 1024]) # 64, 128,
%