### 2020-02-08 code/preparePaper/output/auc_small.tex

Afrikaans & Amharic & Arabic & Armenian & Bambara & Basque & Breton & Bulgarian & Buryat
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Afrikaans-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Amharic-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Arabic-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Armenian-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Bambara-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Basque-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Breton-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Bulgarian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Buryat-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf}
 \\ 
Cantonese & Catalan & Chinese & Croatian & Czech & Danish & Dutch & English & Erzya
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Cantonese-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Catalan-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Chinese-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Croatian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Czech-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Danish-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Dutch-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/English-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Erzya-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf}
 \\ 
Estonian & Faroese & Finnish & French & German & Greek & Hebrew & Hindi & Hungarian
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Estonian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Faroese-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Finnish-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/French-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/German-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Greek-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Hebrew-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Hindi-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Hungarian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf}
 \\ 
Indonesian & Italian & Japanese & Kazakh & Korean & Kurmanji & Latvian & Maltese & Naija
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Indonesian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Italian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Japanese-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Kazakh-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Korean-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Kurmanji-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Latvian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Maltese-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Naija-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf}
 \\ 
North Sami & Norwegian & Persian & Polish & Portuguese & Romanian & Russian & Serbian & Slovak
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/North_Sami-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Norwegian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Persian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Polish-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Portuguese-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Romanian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Russian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Serbian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Slovak-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf}
 \\ 
Slovenian & Spanish & Swedish & Thai & Turkish & Ukrainian & Urdu & Uyghur & Vietnamese
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Slovenian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Spanish-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Swedish-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Thai-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Turkish-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Ukrainian-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Urdu-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Uyghur-Adap-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/medians/auc/figures/Vietnamese-listener-surprisal-memory-HIST_AUC_onlyWordForms_boundedVocab_REAL.pdf}
 \\