### 2020-02-15 code/preparePaper/output/it_small.tex
It plot

Afrikaans & Amharic & Arabic & Armenian & Bambara & Basque & Breton & Bulgarian & Buryat
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Afrikaans-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Amharic-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Arabic-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Armenian-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Bambara-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Basque-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Breton-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Bulgarian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Buryat-Adap-it_REAL.pdf}
 \\ 
Cantonese & Catalan & Chinese & Croatian & Czech & Danish & Dutch & English & Erzya
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Cantonese-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Catalan-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Chinese-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Croatian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Czech-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Danish-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Dutch-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/English-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Erzya-Adap-it_REAL.pdf}
 \\ 
Estonian & Faroese & Finnish & French & German & Greek & Hebrew & Hindi & Hungarian
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Estonian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Faroese-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Finnish-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/French-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/German-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Greek-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Hebrew-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Hindi-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Hungarian-it_REAL.pdf}
 \\ 
Indonesian & Italian & Japanese & Kazakh & Korean & Kurmanji & Latvian & Maltese & Naija
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Indonesian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Italian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Japanese-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Kazakh-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Korean-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Kurmanji-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Latvian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Maltese-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Naija-Adap-it_REAL.pdf}
 \\ 
North Sami & Norwegian & Persian & Polish & Portuguese & Romanian & Russian & Serbian & Slovak
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/North_Sami-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Norwegian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Persian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Polish-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Portuguese-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Romanian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Russian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Serbian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Slovak-it_REAL.pdf}
 \\ 
Slovenian & Spanish & Swedish & Thai & Turkish & Ukrainian & Urdu & Uyghur & Vietnamese
 \\ 
\includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Slovenian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Spanish-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Swedish-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Thai-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Turkish-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Ukrainian-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Urdu-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Uyghur-Adap-it_REAL.pdf} & \includegraphics[width=0.1\textwidth]{../code/analysis/visualize_neural/figures/Vietnamese-it_REAL.pdf}
 \\